Analysis paper

### 2021-10-14 analysis/analysis.Rmd
clean code

--------------------------------------------------------------------------------
## Plot with Visualsiation of daily data

```{r}
df <- copy(dailytruth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
][, target_end_date := as.Date(target_end_date)
]

weeklytruth <- copy(truth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
]

df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  # geom_point(size = 0.3, alpha = 0.4) + 
  # geom_line(size = 0.5, alpha = 0.4) + 
  geom_col(fill = "lightsteelblue", color = "white") + 
  geom_line(data = weeklytruth, aes(y = true_value / 7)) + 
  geom_point(data = weeklytruth, aes(y = true_value / 7), size = 0.3) +
  # geom_vline(aes(xintercept = sundays), linetype = "dashed", alpha = 0.2) + 
  facet_wrap( ~ target_location, scales = "free", ncol = 1) + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                          max(forecast_dates$full_hub_period + 21))) + 
  theme_minimal()

--------------------------------------------------------------------------------
## Plot with Visualsiation of x week ahead forecasts and scores

@@

# create helper functions for the plot 
# ------------------------------------------------------------------------------
# truth_and_pred_plot = helper function for the visualisation truth and forecast
# scores_over_time_plot = helper function for plotting scores over time
# forecasts_and_scores = function that organises creation of the plots

truth_and_pred_plot <- function(df, ttype = "case") {

  if (ttype == "case") {
    dodgewidth = 3.6
  } else {
    dodgewidth = 3.7
  }

  setnames(df, old = c("model"), new = c("Model"))

  trend <- 
    copy(epitrend)[, location_target := paste0(str_to_sentence(target_type), 
                                               "s", " in ", location_name)]
  setnames(trend, old = c("classification"), new = c("Classification"))

  limits <- min(forecast_dates$full_hub_period) + 5 + 1:21 * 7

  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_point(data = trend[target_type == ttype], size = 0.7) +
    geom_line(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  limits <- min(forecast_dates$full_hub_period) + 5 + 1:21 * 7

  p <- scores %>%
    ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
    geom_line(aes(colour = model)) + 
    geom_point(aes(colour = model, shape = classification)) + 
    theme_minimal() + 
    facet_wrap( ~ location_target, ncol = 1, scales = "free") +
    theme(legend.position = "bottom") +
    theme(panel.grid.major.x = element_blank()) +
    scale_y_continuous(labels = scale_fn) + # , 
                       # trans = log_trans()) + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) 
  return(p)
}

forecasts_and_scores <- function(data, 
                                 output_dir = here("analysis", "plots"), 
                                 plotname = "figure-forecasts", 
                                 ensembles = FALSE) {

  if (ensembles) {
    colcases <- c(colcases[1:2], colcases[4], colcases[3])
    coldeaths <- colcases
  }

  # create data.frame for plotting predictions ---------------------------------
  df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

  # create data.frame for plotting scores --------------------------------------
  summarise_by <- c("model", "horizon", "location_name", "target_type", 
                    "forecast_date", "target_end_date", "location_target")

@@

  # copy epitrend data.frame and only keep necessary columns
  epi_tmp <- copy(epitrend)
  epi_tmp[, forecast_date := (as.Date(forecast_date))]
  epi_tmp <- 
    epi_tmp[, .(location_name, forecast_date, classification, target_type)] %>% 
    unique()

  # create plots for all horizons ----------------------------------------------
  for (i in 1:4) {

    # truth plot for cases 
    truth_vs_forecast_cases <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "case") + 
      scale_color_manual(values = colcases) + 
      labs(y = "Forecasts and observed values", x = "")  + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm")) + 
      scale_y_continuous(labels = scale_fn)

    # truth plot for deaths
    truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "death")  + 
      scale_color_manual(values = coldeaths) + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Forecasts and observed values", x = "") +
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm"))

    # scores plot for cases
    scores_over_time_cases <- 
      scores_target_forecastdate[horizon == i & target_type == "case"] %>%
      merge(epi_tmp) %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = colcases) + 
      labs(y = "Weighted interval score", x = "") + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none")

    # scores plot for deaths
    scores_over_time_deaths <- 
      scores_target_forecastdate[horizon == i & target_type == "death"] %>%
      merge(epi_tmp) %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = coldeaths) + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Weighted interval score", x = "")

@@

# create plots -----------------------------------------------------------------
# regular forecasts
unfiltered_data[model %in% c(regular_models)] %>%
  forecasts_and_scores()

data <- unfiltered_data[model %in% c(regular_models, "Renewal-retrospective", "Convolution-retrospective")]

unfiltered_data$model %>% unique()

df$model %>% unique()
# ensemble variants

# models <- c("Hub-ensemble-all", "Crowd forecast", "Hub-ensemble", "Renewal")
# unfiltered_data[model %in% models] %>%
#   forecasts_and_scores(plotname = "figure-forecasts-ensembles", 
#                        ensembles = TRUE)

--------------------------------------------------------------------------------
## Summary tables

@@

# create helper function to make table
# ------------------------------------------------------------------------------
score_table <- function(scores, 
                        group = TRUE, 
                        full_width = TRUE) {
  # copy data, change names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Dispersion", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

  # remove some cols, change column order, sort table and rename horizon
  if (all(c("coverage_deviation", "rel. skill") %in% names(scores))) {
    scores <- scores[, .SD,  .SDcols = !c("coverage_deviation", "rel. skill")]
  }

@@

make_table <- function(data, 
                       plotname = "table_scores",
                       output_dir = here("analysis", "plots")) {

  coverage <- eval_forecasts(
    data = data, 
    summarise_by = c("horizon", "model", "target_type", "range"), 
  )[range %in% c(50, 90), 
    .(model, target_type, coverage, range, horizon)] %>%
    dcast(formula = ... ~ range, value.var = "coverage") %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case"))

  df <- eval_forecasts(
    data, 
    summarise_by = c("model", "horizon", "target_type"), 
    compute_relative_skill = FALSE,
    sd = TRUE) %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")) %>%
    merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

@@

  # no idea how to do that in data.table: 
  df <- df %>%
    dplyr::group_by(horizon, target_type) %>%
    dplyr::mutate_at(percentage_cols, 
                     ~ paste0(., " (", 
                              round(./.[model %in% c("Hub-ensemble", 
                                                     "Hub-ensemble-mean")], 2), 
                              ")"))

@@

  del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                          !(grepl("interval", names(df)))]
  del_cols <- c(del_cols, "coverage_deviation")
  df[, (del_cols) := NULL]

  setnames(df, old = c("interval_score_sd, sharpness"), 
           new = c("WIS - sd", "dispersion"))

  # tables for 1-2 weeks ahead and 3-4 weeks ahead
  table <- score_table(df[horizon <= 2])

  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 2, "_ahead.rds")))

@@

# create tables
# ------------------------------------------------------------------------------

filtered_data[model %in% regular_models] %>%
  make_table()

filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  make_table(plotname = "table-late-period")

filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_mean-ensemble_scores")

filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_median-ensemble_scores")

### Tables for different phases of the epidemic

```{r fig.height=10, fig.width=5}

coverage <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("horizon", "model", "target_type", "target_phase", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, target_phase, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "horizon", "target_type", "target_phase"), 
  metrics = c("interval_score", "coverage_deviation"),
  sd = TRUE, 
  quantiles = 0.5
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_phase", "target_type"))

del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                        !(grepl("interval", names(df)))]
del_cols <- c(del_cols, "coverage_deviation")
df[, (del_cols) := NULL]

setnames(df, old = c("interval_score_0.5", "interval_score_sd"), 
         new = c("WIS - median", 
                 "WIS - sd"))

tables <- list()
table_dfs <- list()
targets <- unique(df$target_phase)

# for horizon <= 2
labels <- df[horizon <= 2][order(target_phase, horizon)]$target_phase
table <- score_table(df[horizon <= 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-2.png"))

# for horizon > 2
labels <- df[horizon > 2][order(target_phase, horizon)]$target_phase
table <- score_table(df[horizon > 2, !"target_type"], group = FALSE) %>%
  pack_rows(index = table(labels), indent = FALSE)
save_kable(table, 
           file = here("analysis", "plots", "table-phases-4.png"))

### Look at ensemble forecasts

#### Number of members in the ensemble
```{r}
ens <- ensemble_members %>%
  melt(id.vars = c("Date", "model"))

@@

ggsave(here("analysis", "plots", "ensemble-members.png"), 
       plot = p)

### Comparing ensembles with and without a model

comparison_ensembles <- function(models_to_plot) {
  df <- unfiltered_data[model %in% models_to_plot &
                          quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

  p <- df[grepl(2, target)] %>%
    ggplot(aes(y = `0.5`, x = target_end_date, color = model)) +
    geom_line() +
    geom_line(aes(y = true_value), color = "black") + 
    theme_minimal() + 
    facet_wrap(~ location_target, scales = "free", ncol = 2) + 
    theme(legend.position = "bottom") + 
    # theme(panel.background = element_rect(fill = "#e1f0fe")) + 
    scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 14, 
                            max(forecast_dates$full_hub_period))) + 
    labs(y = "Median forecast", x = "Date")
  return(p)
}

models_plot <- c("Crowd forecast", "Hub-ensemble-with-crowd", "Hub-ensemble")
p <- comparison_ensembles(models_plot)
ggsave(here("analysis", "plots", "ensemble-with-and-without-crowd.png"), 
       plot = p)

models_plot <- c("Renewal", "Hub-ensemble-with-renewal", "Hub-ensemble")
p <- comparison_ensembles(models_plot)
ggsave(here("analysis", "plots", "ensemble-with-and-without-renewal.png"), 
       plot = p)

models_plot <- c("Renewal", "Crowd forecast", "Hub-ensemble-all", "Hub-ensemble")
p <- comparison_ensembles(models_plot)
ggsave(here("analysis", "plots", "ensemble-with-and-without-both.png"), 
       plot = p)

------------------------------------------------------------------------

## Analyse distribution of scores

```{r}
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

scores[, .(sd = sd(interval_score)), by = c("model", "target_type")]

scores[, target_type := paste0(str_to_title(target_type), "s overall")]

# scores for different phases of the epidemic
scores[, target_phase := factor(target_phase, 
                                levels = str_to_title(c("Cases - decreasing phase", 
                                                        "Cases - unclear phase", 
                                                        "Cases - increasing phase", 
                                                        "Deaths - decreasing phase", 
                                                        "Deaths - unclear phase", 
                                                        "Deaths - increasing phase")))]

distribution_plot_wis <- function(df, x = "interval_score", ranks = FALSE) {

@@

for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score") +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  p_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  p_phases_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
    facet_wrap( ~ target_phase, scales = "free_x")

  # pl_wis <- wrap_plots(p_combined_wis, p_wis, widths = c(1, 2)) / p_phases_wis + 
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, ".png")), 
         plot = pl_wis, height = 6, width = 9)

# plot with rank distributions
for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score", 
                                          ranks = TRUE) +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

@@

  p_phases_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
    facet_wrap( ~ target_phase, scales = "free_x")

  # pl_wis <- wrap_plots(p_combined_wis, p_wis, widths = c(1, 2)) / p_phases_wis + 
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

@@

# same plot for absolute error
p_combined_ae <- distribution_plot_wis(df = scores, 
                                        x = "aem") +
  facet_wrap(~ target_type, ncol = 1, scales = "free_x")

p_ae <- distribution_plot_wis(df = scores, x = "aem") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

```{r}
scores[, .(target_count = .N), by = c("target_phase", "model")]

# p_phases <- distribution_plot(df = scores[model != "Baseline"]) + 
#   facet_wrap( ~ target_phase)

p_phases_ae <- distribution_plot_wis(df = scores, x = "aem") + 
  facet_wrap( ~ target_phase, scales = "free_x")

scores[, .(n = .N), by = "target_phase"]

```{r fig.height=10, fig.width=15}
pl_ae <- wrap_plots(p_combined_ae, p_ae, widths = c(1, 2)) / p_phases_wis + 
  plot_layout(guides = "collect") & 
  plot_annotation(tag_levels = 'A') &
  theme(legend.position = 'bottom')

ggsave("plots/distribution_scores_ae.png", pl_ae, height = 9)

------------------------------------------------------------------------

## WIS contributions

```{r}
# df <- data[grepl("2", target)] %>%
#   merge(epitrend, 
#         by = c("target_end_date", "location_name", "target_type"))

scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification"), 
  compute_relative_skill = TRUE
)

scores_target[, target_phase := paste0(target_type, "s - ", 
                                       classification, " phase")]

wis_components(scores_target, 
               facet_formula = ~ target_phase, 
               scales = "free_x", 
               relative_contributions = TRUE,
               x_text_angle = 0) + 
  theme(legend.position = "bottom") + 
  coord_flip() + 
  labs(x = NULL)

ggsave(here("analysis", "plots", "wis-components.png"))

### Distribution of Bias over time

```{r}
# df <- data[grepl("2", target)] %>%
#   merge(epitrend, 
#         by = c("target_end_date", "location_name", "target_type"))

scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification", "forecast_date")
)

scores_target[, target_phase := paste0(target_type, "s - ", 
                                       classification, " phase")]

p <- scores_target[model %in% regular_models] %>%
  distribution_plot_wis(x = "bias") + 
  scale_x_continuous() + 
  facet_wrap(~ target_phase, ncol = 3) + 
  geom_vline(xintercept = 0, color = "black", size = 1.4) + 
  labs(y = "Bias")

ggsave(here("analysis", "plots", "distribution_bias_phases.png"), plot = p)
```

```{r, eval = TRUE}
scores_full <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "location_target", "horizon", "forecast_date"), 
  compute_relative_skill = FALSE
)

means <- scores_full[, .(meanval = mean(interval_score), 
                         medianval = median(interval_score)), 
                     by =c("model", "location_target", "horizon")]

scores_full %>%
  ggplot(aes(y = interval_score, x = horizon, 
             fill = model, color = model)) +
  geom_boxplot(outlier.shape = 21, coef = 1.5, 
               width=0.5, alpha = 1, 
               position = position_dodge2(width = 0.5)) + 
  # geom_line(data = means, 
  #           aes(group = model, y = meanval, color = model), 
  #           position = position_dodge2(width = 0.5), 
  #           size = 2) +
  # theme(panel.background = element_rect(fill = "#e1f0fe")) +
  scale_color_manual(values = coldeaths) + 
  scale_fill_manual(values = coldeaths) + 
  geom_point(data = means, 
             mapping = aes(y = meanval, fill = model),
             color = "black",
             # fill = "red",
             shape = 21,
             size = 2,
             position = position_dodge2(width = 0.5)) + 
    geom_point(data = means, 
             mapping = aes(y = medianval, fill = model),
             color = "black",
             # fill = "red",
             shape = 22,
             size = 2,
             position = position_dodge2(width = 0.5)) + 
  scale_y_log10() + 
  facet_wrap( ~ location_target, scales = "free") + 
  theme(legend.position = "bottom")

ggsave("plots/scores_horizons.png")

------------------------------------------------------------------------

## Calibration

```{r, eval = TRUE}
scores_target <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "target_type", "classification", "range", "quantile"), 
  compute_relative_skill = FALSE, 
  pit_plots = TRUE
)

interval_coverage(scores_target, 
                  facet_formula = ~ target_type + classification)

```{r, eval = FALSE}
scores <- list()
# scores for restricted data set
scores_full <- eval_forecasts(filtered_data, 
                              summarise_by = c("model", "target_type", "target"), 
                              compute_relative_skill = TRUE)

>>>>

# Plot with Visualsiation of x week ahead forecasts and scores

@@

# helper function which creates panels A and C
truth_and_pred_plot <- function(df, ttype = "case") {

  # adjust distance between bars depending on case or death forecast
  if (ttype == "case") {
    dodgewidth = 3.6
  } else {
    dodgewidth = 3.7
  }

  # rename model column
  setnames(df, old = c("model"), new = c("Model"))

  # define date limits for the plot
  limits <- min(forecast_dates$full_hub_period) + 5 + 0:21 * 7

  # create plot
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = Model), 
                   size = 1.1,
                   alpha = 1,
                   position = position_dodge(width = dodgewidth)) + 
    geom_line(data = unfiltered_data[target_type == ttype & 
                                       target_end_date %in% limits]) + 
    theme_minimal() + 
    theme(panel.grid.major.x = element_blank()) +
    facet_wrap(~ location_target, scales = "free", ncol = 1) + 
    theme(legend.position = "bottom") + 
    scale_x_date(limits = c(min(limits) - 2, max(limits) + 2), 
                 minor_breaks = limits) + 
      scale_y_continuous(labels = scale_fn)
  return(p)
}

# helper function which creates panels B and D
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {

  # define plot limits on x axis
  limits <- min(forecast_dates$full_hub_period) + 5 + 0:20 * 7

  p <- scores %>%
    ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
    geom_line(aes(colour = model)) + 
    geom_point(aes(colour = model)) + 
    theme_minimal() + 
    facet_wrap( ~ location_target, ncol = 1, scales = "free") +
    theme(legend.position = "bottom") +
    theme(panel.grid.major.x = element_blank()) +
    scale_y_continuous(labels = scale_fn) + # , 
                       # trans = log_trans()) + 
    scale_x_date(limits = c(min(limits), max(limits)), 
                 minor_breaks = limits) 
  return(p)
}

# helper function that creates the final plot
forecasts_and_scores <- function(data, 
                                 output_dir = here("analysis", "plots"), 
                                 plotname = "figure-forecasts") {

  # calculate scores
  summarise_by <- c("model", "horizon", "location_name", "target_type", 
                    "forecast_date", "target_end_date", "location_target")

@@

  # create data.frame for plotting predictions 
  df <- data[quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")

  # create plots for all horizons 
  for (i in 1:4) {

    # truth plot for cases 
    truth_vs_forecast_cases <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "case") + 
      scale_color_manual(values = colcases, name = "Model") + 
      labs(y = "Forecasts and observed values", x = "")  + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm")) + 
      scale_y_continuous(labels = scale_fn)

    # truth plot for deaths
    truth_vs_forecast_deaths <- df[grepl(i, target)] %>%
      truth_and_pred_plot(ttype = "death")  + 
      scale_color_manual(values = coldeaths, name = "Model") + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 5 + (i-1) * 7,
               xmax = max(forecast_dates$death_not_scored) + 5 + (i-1) * 7,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Forecasts and observed values", x = "") +
      theme(legend.position = "none") +
      guides(color = "none", shape = "none") + 
      theme(plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm"))

    # scores plot for cases
    scores_over_time_cases <- 
      scores_target_forecastdate[horizon == i & target_type == "case"] %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = colcases, name = "Model") + 
      labs(y = "Weighted interval score", x = "") + 
      theme(legend.position = "none") +
      guides(color = "none", shape = "none")

    # scores plot for deaths
    scores_over_time_deaths <- 
      scores_target_forecastdate[horizon == i & target_type == "death"] %>%
      scores_over_time_plot(y = "interval_score") + 
      scale_color_manual(values = coldeaths, name = "Model") + 
      annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
               xmax = max(forecast_dates$death_not_scored) + 14,
               ymin = 0,
               ymax = Inf, fill = "grey10", alpha = 0.1) + 
      labs(y = "Weighted interval score", x = "")

@@

# create plots -----------------------------------------------------------------
# regular forecasts
unfiltered_data[model %in% c(regular_models)] %>%
  forecasts_and_scores()
```

# Create tables with scores

@@

# helper function to create the tables
score_table <- function(scores, 
                        group = TRUE, 
                        full_width = TRUE) {

  # change column names, round column values and reduce to sign. digits
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("horizon", "model", "target_type", "interval_score", "aem", 
                   "relative_skill", "scaled_rel_skill", "sharpness", 
                   "underprediction", "overprediction", "bias", 
                   "50", "90", "location_target", "target_phase"), 
           new = c("Horizon", "Model", "Target", "WIS", "Abs. error", 
                   "rel. skill", "WIS - rel.", "Dispersion", 
                   "Underpred.", "Overpred.", "Bias", 
                   "50%-Cov.", "90%-Cov.", "Target", "Target"), 
           skip_absent = TRUE)

  # remove some cols, change column order, sort table
  if (all(c("coverage_deviation", "rel. skill") %in% names(scores))) {
    scores <- scores[, .SD,  .SDcols = !c("coverage_deviation", "rel. skill")]
  }

@@

# function that actually makes the tables
make_table <- function(data, 
                       plotname = "table_scores",
                       output_dir = here("analysis", "plots")) {

  # get coverage values
  coverage <- eval_forecasts(
    data = data, 
    summarise_by = c("horizon", "model", "target_type", "range"), 
  )[range %in% c(50, 90), 
    .(model, target_type, coverage, range, horizon)] %>%
    dcast(formula = ... ~ range, value.var = "coverage") %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case"))

  # compute scores and merge with coverage values
  df <- eval_forecasts(
    data, 
    summarise_by = c("model", "horizon", "target_type"), 
    compute_relative_skill = FALSE,
    sd = TRUE) %>%
    filter(!(model %in% c("Hub-ensemble-with-all", 
                          "Hub-ensemble-with-convolution", 
                          "Hub-ensemble-with-all-mean", 
                          "Hub-ensemble-with-convolution-mean") & 
               target_type == "case")) %>%
    merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

@@

  # add percentage values in brackets
  df <- df %>%
    dplyr::group_by(horizon, target_type) %>%
    dplyr::mutate_at(percentage_cols, 
                     ~ paste0(., " (", 
                              round(./.[model %in% c("Hub-ensemble", 
                                                     "Hub-ensemble-mean")], 2), 
                              ")"))

@@

  # delete and rename a few columns
  del_cols <- names(df)[(grepl("_sd", names(df)) | grepl("0.5", names(df))) &
                          !(grepl("interval", names(df)))]
  del_cols <- c(del_cols, "coverage_deviation")
  df[, (del_cols) := NULL] 
  setnames(df, old = c("interval_score_sd, sharpness"), 
           new = c("WIS - sd", "dispersion"))

  # create and save tables for 1-2 weeks ahead and 3-4 weeks ahead
  table <- score_table(df[horizon <= 2]) 
  saveRDS(object = table, 
          file = here("analysis", "plots", paste0(plotname, "_", 2, "_ahead.rds")))

@@

# create tables

# regular models
filtered_data[model %in% regular_models] %>%
  make_table()

# regular models with later period as sensitivity analysis in SI
filtered_data[model %in% regular_models] %>%
  filter(!(forecast_date %in% as.character(forecast_dates$death_not_scored))) %>%
  make_table(plotname = "table-late-period")

# table for mean ensembles for SI
filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_mean-ensemble_scores")

# table for median ensembles for SI
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  make_table(plotname = "table_median-ensemble_scores")
```

# Plot with Visualsiation of daily data

# reformat daily truth data
df <- copy(dailytruth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
][, target_end_date := as.Date(target_end_date)
]

# reformat weekly truth data
weeklytruth <- copy(truth_data)[
  , target_location := paste0(str_to_title(target_type), "s in ", location_name)
]

# create plot
df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_col(fill = "lightsteelblue", color = "white") + 
  geom_line(data = weeklytruth, aes(y = true_value / 7)) + 
  geom_point(data = weeklytruth, aes(y = true_value / 7), size = 0.3) +
  # geom_vline(aes(xintercept = sundays), linetype = "dashed", alpha = 0.2) + 
  facet_wrap( ~ target_location, scales = "free", ncol = 1) + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                          max(forecast_dates$full_hub_period + 21))) + 
  theme_minimal()

# Number of member models in the Hub ensemble

```{r}
ens <- ensemble_members %>%
  melt(id.vars = c("Date", "model"))

@@

ggsave(here("analysis", "plots", "ensemble-members.png"), 
       plot = p)
```

# Plot with distribution of scores

```{r}
# score forecasts
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

# reformat column
scores[, target_type := paste0(str_to_title(target_type), "s overall")]

# display standard deviation for different models
scores[, .(sd = sd(interval_score)), by = c("model", "target_type")]

# helper function to create distribution plot
distribution_plot_wis <- function(df, x = "interval_score", 
                                  ranks = FALSE) {

@@

# create plots for the distributions of WIS
for (i in 1:4) {
  # plot for locations combined
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score") +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

  # plot for locations separate
  p_wis <- distribution_plot_wis(df = scores[horizon == i]) + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y = element_blank()) + 
  facet_wrap(~ location_target, ncol = 2, scale = "free")

  # combine plots and save
  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

  ggsave(here("analysis", "plots", 
              paste0("distribution_scores_wis-", i, ".png")), 
         plot = pl_wis, height = 6, width = 9)
}

# create plots for the distributions of WIS rank
for (i in 1:4) {
  p_combined_wis <- distribution_plot_wis(df = scores[horizon == i], 
                                          x = "interval_score", 
                                          ranks = TRUE) +
    facet_wrap(~ target_type, ncol = 1, scales = "free") + 
    theme(plot.margin=unit(c(0,0.4,0,0),"cm")) + 
    theme(legend.position = "none") + 
    guides(shape = "none", colour = "none", fill = "none")

@@

  pl_wis <- p_combined_wis + p_wis +
    plot_layout(guides = "collect", 
                widths = c(1, 2)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

@@

}
```

# Number of expert forecasters
